TacticToe: Learning to Prove with Tactics

### 2018-02-20 tactictoe_jv.tex
minor

\paragraph{Goals and theorems.}
A goal is represented as a \emph{sequent}.
A sequent is composed of a set of assumptions and a conclusion, all of which are higher-order logic formulas.
When a goal is proven, it becomes a theorem.
If the developer gives the theorem a name then we refer to it as a top-level theorem.

@@

\paragraph{Tactics in \holfour.}
Considering concrete implementation details now, a \emph{tactic} is a function
that takes a goal (or proof obligation) and returns a list of goals (subgoals
of the original goal) together
with a validation function.
Calling the validation function on a list of \emph{theorems}, corresponding to
the list of subgoals, results in a theorem corresponding to the original goal.
For example, if the list of subgoals is empty, then calling the validation
function on the empty list yields the originally goal sequent as a theorem.
In this way, tactics implement the plumbing of backwards, goal-directed, proof
construction.
Because validation functions are only important to check the final proof, we
omit them during the description of proof search algorithm. We denote by $t(g)$
the list of goals produced by the application of a tactic $t$ to a goal $g$.

@@

The application of a tactic to an initial conjecture produces a list of goals.
Each of the created goals can in turn be the input to other tactics.
Moreover, it is possible to try multiple tactics on the same goal.
In order to keep track of the progress made so far from the initial conjecture we want to solve, we organize goals and tactics into a graph with list of goals as nodes and tactics as edges (See Definition~\ref{def:stree}).

In this section, we only give a description of the graph at a given
moment of the search after a certain amount of tactic applications.
Exploration of the tree is directed by the MCTS algorithm in
Section~\ref{sec:proofsearch}, which is guided by prediction algorithms
presented in Section~\ref{sec:predictions}.

@@

In the following, properties are defined in the context of a
search tree $\mathfrak{T}$. In Definition~\ref{def:solved}, we specify
different status for a goal: open, pending or solved. There, we define
what a solved goal and solved node are by mutual recursion on the number of
steps it takes to solve a goal and a node.
\begin{definition}\label{def:solved}(solved goal, solved nodes, open goal,
pending goal)\\
The set of solved nodes $\mathbb{A}^\infty$ and
the set of solved goals $\mathbb{G}^\infty$ are defined inductively by:

@@

We call a goal (respectively node) unsolved if it is not in the set of solved
goals (respectively nodes).
The open goal of a node is the first unsolved goal of this node according to
some preset order. The other unsolved nodes are called pending nodes.
If a node is unsolved, then it contains a unique open goal.
During MCTS exploration, we do not explore pending goals of an unsolved node
before its open goal is solved which justifies our terminology. In other words,
our search tree has the following property: if $g$ is a pending node then
$T(g)$ is empty.

@@

In Definition~\ref{def:prod}, we decide if a tactic is productive based on its
contribution to the search tree.

\begin{definition}\label{def:prod} (productive tactic)
The application of a tactic on an open goal is called productive if and only if
all the following conditions are satisfied:
\begin{itemize}
\item It does not fail or timeout. To prevent tactics from looping, we
interrupt tactics after a short amount of time (0.05 seconds).
\item It does not loop, i.e., its output does not contain any goals that appear
in the ancestor of the node of the current goal.
\item It is not a parallel step, i.e., its output is not a superset
of a list of goals produced from the same goal.
\end{itemize}

@@

\section{Prediction}\label{s:prediction}
The learning-based selection of relevant lemmas significantly improves the
automation for hammers~\cite{BlanchetteGKKU16}. In \tactictoe we use the
distance-weighted \emph{$k$ nearest-neighbour} classifier~\cite{DudaniS76}
adapted for lemma selection~\cite{ckju-pxtp13}. It allows for hundreds of
predictions per second, while maintaining very high prediction quality~\cite{femalecop}.
We rely on this supervised machine learning algorithm to predict three
different kinds of objects:
 tactics, theorems and lists of goals. The number of predicted objects $k$ is
 called the \emph{prediction radius}. For a goal $g$, we note
$\mathit{Predictions}^{\mathit{tactic}}_k (g)$ (respectively
$\mathit{Predictions}^{\mathit{theorem}}_k (g)$
and $\mathit{Predictions}^{\mathit{goal\_list}}_k (g)$) the $k$ tactics
(respectively theorems and lists of goals) selected by our prediction algorithm.

We discuss below the specifics associated with the prediction of each kind of object.
In particular, we present the dataset from which objects are selected and the
purpose of the selection process. The similarity measure backing the
predictions is described in Section~\ref{sec:features} and two methods for
improving the quality of the predictions are presented in
Section~\ref{sec:ortho} and Section~\ref{sec:synthesis}.

\paragraph{Tactics}
To start with, we build a database of tactics consisting
of tactic-goal pairs recorded from successful tactic applications in human
proofs (recording will be discussed in 
Section~\ref{sec:recording}).
During proof search, the recorded goals will be reordered according to their
similarity with a target goal $g$ (usually the first pending goal of a node).
The recorded pairs induce an
order on tactics. The intuition is that tactics which have been successful on
goals similar to $g$ should be tried first, since they are more likely to lead
to a proof.
This predicted tactic order is translated into a prior policy for our MCTS
algorithm (see Section~\ref{sec:policy}).
Tactic selection is also used to improve the quality of the database
of tactics during orthogonalization (see Section~\ref{sec:ortho}).

@@

\paragraph{Lists of Goals as Output of Tactics}
A dataset of tactic outputs is compiled during orthogonalization.
This set is separated into positive and negative examples (see
Section~\ref{sec:evaluation}).
List of goals that are a subset of the result of a recorded tactic on a
target goal are considered positive. During proof search, given a list of
goals $l$ created by a tactic, we select a set of lists of goals that are most
similar to $l$. The ratio of positive examples in the selection gives us a
prior evaluation for MCTS.

@@

Despite the best efforts of the prediction algorithms, a selected tactic may
not solve the current goal, proceed in the wrong direction
or even loop. For that reason, predictions need to be accompanied by a
proof search mechanism that allows for backtracking and
can choose which proof tree to extend next and in which direction.
Our search algorithm is a Monte Carlo tree search~\cite{montecarlo} algorithm
that relies
on a prior evaluation function and a prior policy function. Those priors are
learned through direct supervised learning for the policy and via the data
accumulated during orthogonalization for the evaluation. Therefore, this MCTS
algorithm does not need to rely on roll outs for evaluation. The
policy and evaluation are estimated by the simple supervised k-NN algorithm
here instead of neural networks trained through reinforcement learning in
AlphaGo Zero~\cite{silver2017mastering}.
A characteristic of the MCTS algorithm is that it offers a good trade-off
between exploitation and exploration. This means that the algorithm searches
deeper more promising branches and leaves enough time for the exploration
of less likely alternatives.

@@

\paragraph{Initialization}
The input of the algorithm is a goal $g$ (also called conjecture) that we
want to prove.
\todo{@TG: Check sentence below}
Therefore the search tree starts with only one node the list of goals
$[g]$ and an ordered list of predicted tactics for the goal $g$.
%Then, the MCTS algorithm alternates node selection steps and node extension
%steps which grow the search tree.

%A similar trade-off between exploration and exploitation exists in algorithm
%like Monte Carlo. However, Monte Carlo tree search metrics are based on many
%simulations and do not reflect our exploration strategy that only
\paragraph{Node selection}
Through node extension steps the search tree grows and the number of paths to
explore increases. To decide which node to extend next, an evolving value for
each node is given in Definition~\ref{def:value}. The algorithm that performs
node selection starts from the root of the search tree. From the current node,
the function \texttt{CompareChildren} choses among the children of its open
goal the one with the highest value.

@@

\begin{definition}\label{def:value}(Value)
The value of the $i^{th}$ child $a_i$ of the open goal of a parent node $p$
is determined by:
\[\mathit{Value}(a_i) = \mathit{CurEvaluation}(a_i) + c_{exploration} *
\mathit{Exploration}(a_i)\]

@@

\[\mathit{CurPolicy}(a_i) = \frac{1 +
\mathit{Visit}(a_i))}{\sqrt{\mathit{Visit}(p))}}\]
The policy is approximately the  percentage of time a node was visited. The
square root skews this probability to favor even more exploration of nodes with
few visits.

@@

\paragraph{Backpropagation}
During backpropagation we update the statistics of all the nodes traversed or
created during this MCTS step:
\begin{itemize}
\item Their visit number is incremented.
\item If node extension failed, their failure count is incremented.
\item If node extension succeeded, they inherit the evaluation of the created
child.
\end{itemize}

These changes update the current evaluation and the current policy of the
traversed nodes.
After completing backpropagation, the process loops back to node
selection.

\paragraph{Resolution}
The search ends when the algorithm reaches one of these 3 conditions:
\begin{itemize}
\item It finds a proof, i.e., the root node is solved. In this case,
the search returns a minimized and
prettified proof script (see Section~\ref{sec:proofdisplay}).
\item It saturates, i.e., there are no tactics to be applied to any open goal.
This happens less than 10 times in the full-scale experiment. \todo{@TG: why don't we have the original tactics as options?}
\item It exceeds a given time limit (default is set to 60 seconds).
\end{itemize}

\subsection{Supervised learning guidance}
Our MCTS algorithm relies on the guidance of two supervised learning methods.
The prior policy decides the order in which tactic should be tried and the
prior evaluation estimates if a branch is fruitful by analysing tactic
outputs. Both priors influence the rates at which branches of the search tree
are explored.

\paragraph{Prior policy}\label{sec:policy}
Co-distances between goals produced by the tactic predictor are hard
to translate into a prior policy, so we rely solely on the induced order of
tactics to estimate the probability with which each tactic should be tried.
Let $\mathfrak{T}$ be a search tree after a number of MCTS steps.
Let $p$ be a selected node in $\mathfrak{T}$ and $g$ its open goal. We order
the list of $n$ productive tactics already
applied to $g$ by their similarity score so that $T(g)= \lbrace
t_0,\ldots,t_{n-1} \rbrace$.
Let $c_{policy}$ be a constant heuristic that estimates how likely a tactic
$t_i$ is to prove the goal.
We can now calculate the prior policy of a child $a_i$ produced by the tactic
$t_i$ by:
\[PriorPolicy(a_i) = (1 - c_{policy})^{i} * c_{policy}\]

@@

We now concentrate on the definition of a reasonable evaluation function for the
output of tactics. Theoretically, a list of goals is interesting if short proofs
are possible each the goals in the list, and the lengths of the proofs can be
estimated from previous proof searches. However, this
approach generates too much data and slows down the proof search. So, we only
collect outputs of tactics tested during orthogonalization. We
declare a list of goals $l$ to be positive if it is has been produced by the
winner of an orthogonalization competition.
The set of positive examples in
$\mathit{Predictions}^{\mathit{goal\_list}}_k(l)$ is noted
$\mathit{Positive}^{\mathit{goal\_list}}_k(l)$.
We chose $k=10$ as a default evaluation radius in our experiments.
And we evaluate a node $a$ through the list of goals $G(a)$ it
contains using the prior evaluation function:
\begin{align*}
\mathit{PriorEvaluation}_k (a) &=
  \frac{card\ \mathit{Positive}^{\mathit{goal\_list}}_k(G(a)}{k}\\
\end{align*}

\subsection{ATP Integration}\label{sec:atp}
General-purpose proof automation mechanisms which combine proof translation to
ATPs with machine learning (``hammers'') have become quite successful in
enhancing the automation level in proof assistants~\cite{hammers4qed}.
Since external automated reasoning techniques sometimes outperform the combined
power of tactics, we would like to combine the \tactictoe search with
general purpose provers such as the ones found in \holyhammer for
\holfour~\cite{tgck-cpp15}.

\holfour already integrates a superposition-based prover \metis for this purpose. It is already
recognized by the tactic selection mechanism in \tactictoe and its
premises can be predicted dynamically when in an abstracted form. Nevertheless,
we think that the performance of \tactictoe can be boosted by giving the ATP
\metis a special status. This arrangement consists of always predicting
premises for \metis, giving it a slightly higher timeout and
trying it first on each open goal. These modifications only induce an
linear overhead on the number nodes as \metis does not create
new goals.

The integration of an external prover is also possible. For this purpose we experiment
with 
\eprover~\cite{eprover}.
Namely, we developed a tactic which expects a list of premises as an
argument
and calls \eprover on the goal together with the premises translated to a first-order problem. If \eprover
succeeds on the problem, the premises used in the external proof are used to
reconstruct the proof inside \holfour with \metis. Giving a special status
to external
provers is essential as external ATP calls do not appear in human
proof scripts. For external ATPs, we use an even higher timeout (5 seconds) and
a larger number of predicted premises (128 for \eprover) and also try them
first on open
goals. Since calls to ATPs are
computationally expensive, they are run in parallel and asynchronously. This
avoids slowing down \tactictoe's search loop. The number of asynchronous calls
to external provers that can be executed at the same time is limited by the
number of cores available to the user.

@@

Here are the different kind of feature we extract from terms:
\begin{itemize}
\item names of constants, including the logical operators,
\item type constructors present in the types of constants and variables,
\item subterms with all variables replaced by a single placeholder $V$,
\item names of the variables.
\end{itemize}
Goals (respectively theorems) are represented by a couple of a list of
terms (assumptions) and a term (conclusion). So, we distinguish between
features of the assumptions and features of the conclusion by adding a
different tag to elements of each set. The union of all those
features are considered the feature of a goal (respectively theorem). From
that, we can construct features for a list of goals by
from the union of the features of each goal in the list.

@@

The first idea is to select tactic-goal pairs and theorems by their similarity
with $c$ (i.e. $\mathit{Predictions}^{\mathit{tactic}}_{500} (c)$ and
$\mathit{Predictions}^{\mathit{theorem}}_{500} (c)$).
However, this selection may not be adapted to later stages of
the proof where goals may have drifted significantly from $c$. In order to
anticipate the production of diverse goals, our prediction during preselection
takes dependencies between objects of the same dataset into account.
This dependency relation is assymetric. Through the relation, each object has
multiple children and at most one parent.
Once a dependency relation is established we can calculate a dependency score
for each object, which is the maximum of its similarity score
and the similarity score of its parent. In the end, the 500 entries with
highest dependency score are preselected in each dataset.

@@

\section{Abstraction}\label{sec:synthesis}
One of the major weakness of the previous version of \tactictoe was that
it could not create its own tactics. Indeed, sometimes no previous tactic is
adapted for a certain goal and creating a new tactic is necessary.
%Experiments in section, show how different degrees of tactic synthesis could
%improve the success rate of \tactictoe.
In this section we present a way to create tactics with different arguments
by abstracting them and re-instantiating them using a predictor for that kind
of argument. In the spirit of the orthogonalization method, we will try to
create tactics that are more general but have the same effect as the
ungeneralized one. The generalized tactics are typically slower that their
original variants, but the added flexibility is worthwhile in practice. Moreover,
since we impose a timeout on each tactic (0.05 seconds by default), very slow
tactics fail and thus are not selected.

\paragraph{Abstraction of Tactic Arguments}
The first step is to abstract arguments of tactics by replacing them by a
placeholder, creating a tactic with an unspecified argument.
Suppose we have a recorded tactic $t$. Since $t$ is a \sml code tree, we can
try to abstract any of the
\sml subterms. Let $u$ be a \sml subterm of $t$, and $h$ a placeholder in $u$, we
can substitute $u$ by $h$ to create an abstracted tactic.
We denote this intermediate tactic $t[h/u]$. By repeating the
process, multiple
arguments may be abstracted in the same tactic. Ultimately, the more
abstractions are performed the more general a tactic is, as many tactics
become instances of the same abstracted tactic. As a consequence, it becomes
harder and harder to predict suitable arguments. In our experiments, we create
one abstraction $\hat{t}$ for each tactic $t$ by abstracting all subterms of
type theorem list in $t$.

\paragraph{Instantiation of an Abstracted Tactic}
An abstracted tactic is not immediately applicable to a goal, since it contains
unspecified arguments. To apply an abstracted tactic $\hat{t}$, we first need
to
instantiate the placeholders inside $\hat{t}$. Because it is difficult to
build a general predictor effective on each type of argument, we manually
design different argument predictors for each type. Those predictors are
given a goal $g$ as input and return the best predicted argument for this
goal.

Our default algorithm relies on argument predictions for theorems with a
radius of 16. It uses the produced list
$\mathit{Predictions}^{\mathit{tactic}}_{16}(g)$ to replace all the
placeholders
in $\hat{t}$.
The type of theorem lists is a very common type in \holfour proofs and theorems contain
enough information to be predicted accurately.
We also have positive results with term abstraction and plan to include it soon
in our default algorithm.

\paragraph{Selection of Abstracted Tactics}
As abstracted tactics do not appear in human proofs, we need to find a way to
predict them so that they can contribute to the \tactictoe proof search.
A straightforward idea is to try $\hat{t}$ before $t$ during the proof search.
However, this risks doing unnecessary work as the two may perform similar steps.
Therefore, we would like to decide
beforehand if one is better than the other.
In fact, we can re-use the orthogonalization module to do this task for us.
We add $\hat{t}$ to the competition, initially giving it the coverage of $t$.
If $\hat{t}$
wins, it is associated with $g$ and is included in the database of tactic
features and thus can be predicted during proof search.
After many orthogonalization competitions, the coverage of $\hat{t}$ may exceed
the coverage of
$t$. At
this point, the coverage of $\hat{t}$ is estimated on its own and not inherited
from
$t$ anymore.

@@

\subsection{Tuning \tactictoe}\label{sec:tuning}

@@

%(more than 10 steps long). In all those experiments orthogonalization is

@@

\subsection{Full-scale experiment}\label{sec:full_exp}

@@

Table~\ref{theories} compares the re-proving success rates for different
\holfour theories. \tactictoe outperforms \eprover on every
considered theory.
\eprover is more suited to deal with dense theories such as
\texttt{real} or \texttt{complex} where a lot of related theorems are available
and most proofs are usually completed by rewriting tactics. Thanks to its
ability to re-use custom-built tactics, \tactictoe
largely surpasses \eprover on theories relying on inductive terms and
simplification sets such as \texttt{arithmetic}, \texttt{list}
and \texttt{f\_map}. Indeed, \tactictoe is able to recognize where and when to
apply induction, a task at which ATPs are known to struggle with.

>>>>

\paragraph{Goals and theorems.}
A goal is represented as a sequent. Each sequent is composed of
a list of assumptions and a conclusion which are higher-order logic formulas.
When a goal is proven, it becomes a theorem. And if the developer gives it a
name then we refer to it as a top-level theorem.

@@

\paragraph{Tactics in \holfour.}
Considering concrete implementation details now, a \emph{tactic} is a function
that takes a goal (or proof obligation) and returns a list of goals (subgoals 
of the original goal) together
with a validation function.
Calling the validation function on a list of \emph{theorems}, corresponding to 
the list of subgoals, results in a theorem corresponding to the original goal.
For example, if the list of subgoals is empty, then calling the validation 
function on the empty list yields the originally goal sequent as a theorem.
In this way, tactics implement the plumbing of backwards, goal-directed, proof 
construction. 
Because validation functions are only important to check the final proof, we 
omit them during the description of proof search algorithm. We note $t(g)$ 
the lists of goals produced by the application of a tactic $t$ to a goal 
$g$.

@@

The application of tactic to an initial conjecture produces a list of goals. 
Each of the created goals can in turn be the input other tactics. Moreover, it 
is possible to try multiple tactics on the same goal. In order to keep track of 
the progress made so far from the initial conjecture we want to solve, we 
organize goals and tactics into a graph with list of goals as nodes and tactics 
as edges (See Definition~\ref{def:stree}).

In this section, we only give a description of the graph at a given 
moment of the search after a certain amount of tactic applications. 
The exploration of the tree is decided by MCTS algorithm in 
Section~\ref{sec:proofsearch}, which is guided by a prediction algorithms 
presented in Section~\ref{sec:predictions}.

@@

In the following, properties are defined in the context of a
search tree $\mathfrak{T}$. In Definition~\ref{def:solved}, we specify 
different status for a goal: open, pending or solved. There, we define 
what a solved goal and solved node are by mutual recursion on the number of
steps it takes to solve a goal and a node.
\begin{definition}\label{def:solved}(Solved goal, solved nodes, open goal,
pending goal)\\
The set of solved nodes $\mathbb{A}^\infty$ and
the set of solved goals $\mathbb{G}^\infty$ are defined inductively by:

@@

We call a goal (respectively node) unsolved if it is not in the set of solved  
goal (respectively node).
The open goal of a node is the first unsolved goal of this node according to 
some preset order. The other unsolved nodes are called pending nodes.
If a node is unsolved, then it contains a unique open goal.
During MCTS exploration, we do not explore pending goals of an unsolved node 
before its open goal is solved which justifies our terminology. In other words, 
our search tree has the following property: if $g$ is a pending node then 
$T(g)$ is empty.

@@

In Definition~\ref{def:prod}, we judge if a tactic is productive from its 
contribution to the search tree.

\begin{definition}\label{def:prod} (Productive tactic)
The application of a tactic on an open goal is called productive if and only if
all the following conditions are satisfied:
\begin{itemize}
\item It does not fail or timeout. To prevent tactics from looping, we
interrupt tactics after a short amount of time (0.05 seconds).
\item It does not loop, i.e. its output does not contain any goals that appears 
in the ancestor of the node of the current goal.
\item It is not a parallel step, i.e. its output is not a superset
of a list of goals produced from the same goal.
\end{itemize}

@@

\section{Prediction}\label{s:prediction}
The learning-based selection of relevant lemmas significantly improves the
automation for hammers~\cite{BlanchetteGKKU16}. In \tactictoe, we use the
distance-weighted \emph{$k$ nearest-neighbour} classifier~\cite{DudaniS76}
adapted for lemma selection~\cite{ckju-pxtp13}. It allows for hundreds of
predictions per second, while maintaining very high prediction quality~\cite{femalecop}.
We rely on this supervised machine learning algorithm to predict three 
different kinds of objects:
 tactics, theorems and lists of goals. The number of predicted objects $k$ is
 called the \emph{prediction radius}. For a goal $g$, we note
$\mathit{Predictions}^{\mathit{tactic}}_k (g)$ (respectively
$\mathit{Predictions}^{\mathit{theorem}}_k (g)$
and $\mathit{Predictions}^{\mathit{goal\_list}}_k (g)$) the $k$ tactics 
(respectively theorems and lists of goals) selected by our prediction algorithm.

We introduce specificities associated with the prediction of each object.
In particular, we present the dataset from which objects are selected and the
purpose of the selection process. Details on the similarity measure backing the 
predictions is described in Section~\ref{sec:features} and two methods for 
improving the quality of the predictions are presented in 
Section~\ref{sec:ortho} and Section~\ref{sec:synthesis}.

\paragraph{Tactics}
To start with, we build a database of tactics consisting
of tactic-goal pairs recorded from successful tactic applications in human
proofs (see
Section~\ref{sec:recording}).
Then, during proof search, we order recorded goals according to their
similarity with a targeted goal $g$ (usually the first pending goal of a node).
The recorded pairs induce an
order on tactics. The intuition is that tactics which have been successful on
goals similar to $g$ should be tried first, since they are more likely to lead
to a proof.
This predicted tactic order is translated into a prior policy for our MCTS
algorithm (see Section~\ref{sec:policy}).
Tactic selection is also used to improve the quality of the database
of tactics during orthogonalization (see Section~\ref{sec:ortho}).

@@

\paragraph{Lists of Goals as Output of Tactics}
A dataset of tactic outputs is compiled during orthogonalization.
This set is separated into positive and negative examples (see
Section~\ref{sec:evaluation}).
List of goals that are a subset of the result of a recorded tactic on a
targeted goal are considered positive. During proof search, given a list of
goals $l$ created by a tactic, we select a set of lists of goals that are most
similar to $l$. The ratio of positive examples in the selection gives us a
prior evaluation for MCTS.

@@

Despite the best efforts of the prediction algorithms, a selected tactic may
not solve the current goal, proceed in the wrong direction
or even loop. For that reason, predictions need to be accompanied by a
proof search mechanism that allows for backtrac king and
can choose which proof tree to extend next and in which direction.
Our search algorithm is a Monte Carlo tree search~\cite{montecarlo} algorithm 
that relies
on a prior evaluation function and a prior policy function. Those priors are
learned through direct supervised learning for the policy and via the data
accumulated during orthogonalization for the evaluation. Therefore, this MCTS
algorithm does not need to rely on roll outs for evaluation. The
policy and evaluation are estimated by the simple supervised k-NN algorithm
here instead of neural networks trained through reinforcement learning in
AlphaGo Zero~\cite{silver2017mastering}.
A characteristic of the MCTS algorithm is that it offers a good trade-off
between exploitation and exploration. This means that the algorithm searches
deeper more promises branches and leaves enough time for the exploration
of less likely alternatives.

@@

\paragraph{Initialization}
The input of the algorithm is a goal $g$ (also called conjecture) that we 
want to prove.
Therefore the search tree starts with only one node the list of goals
$[g]$ and an ordered list of predicted tactics for the goal $g$.
%Then, the MCTS algorithm alternates node selection steps and node extension
%steps which grow the search tree.

%A similar trade-off between exploration and exploitation exists in algorithm
%like Monte Carlo. However, Monte Carlo tree search metrics are based on many
%simulations and do not reflect our exploration strategy that only
\paragraph{Node selection}
Through node extension steps the search tree grows and the number of paths to
explore increases. To decide which node to extend next, a evolving value for
each node is given in Definition~\ref{def:value}. The algorithm that performs
node selection starts from the root of the search tree. From the current node,
the function \texttt{CompareChildren} choses among the children of its open 
goal the one with the highest value.

@@

\begin{definition}\label{def:value}(Value)
The value of the $i^th$ child $a_i$ of the open goal of a parent node $p$ 
is determined by:
\[\mathit{Value}(a_i) = \mathit{CurEvaluation}(a_i) + c_{exploration} * 
\mathit{Exploration}(a_i)\]

@@

\[\mathit{CurPolicy}(a_i) = \frac{1 + 
\mathit{Visit}(a_i))}{\sqrt{\mathit{Visit}(p))}}\]
The policy is approximatively the  percentage of time a node was visited.The
square root skews this probability to favor even more exploration of nodes with
few visits.

@@

\paragraph{Backpropagation}
During backpropagation we update the statisitics of all the nodes traversed or
created during this MCTS step:
\begin{itemize}
\item Their number of visit is incremented.
\item If node extension failed, their failure count is incremented.
\item If node extension succeeded, they inherit the evaluation of the created
child.
\end{itemize}

These changes updates the current evaluation and the current policy of the
traversed nodes.
After completing backpropagation, the process loops back to node
selection.

\paragraph{Resolution}
The search ends when the algorithm reaches one of these 3 conditions:
\begin{itemize}
\item It finds a proof, i.e. the root node is solved. In this case,
the search returns a minimized and
prettified proof script (see Section~\ref{sec:proofdisplay}).
\item It saturates, i.e. there is no tactics to be applied to any open goal.
This happens less than 10 times in the full-scale experiment.
\item It exceeds a given time limit set by default to 60 seconds.
\end{itemize}

\subsection{Supervised learning guidance}
Our MCTS algorithm relies on the guidance of two supervised learning methods,
the prior policy decides the order in which tactic should be tried and the 
prior evaluation estimates if a branch is fruitful by analysing tactic 
outputs. Both priors influence the rates at which branches of the search tree 
are explored.

\paragraph{Prior policy}\label{sec:policy}
Co-distances between goals produced by the tactic predictor are hard
to translate into a prior policy, so we rely solely on the induced order of
tactics to estimate the probability with which each tactic should be tried.
Let $\mathfrak{T}$ be a search tree after a number of MCTS steps.
Let $p$ be a selected node in $\mathfrak{T}$ and $g$ its open goal. We order 
the list of $n$ productive tactics already
applied to $g$ by their similarity score so that $T(g)= \lbrace
t_0,\ldots,t_{n-1} \rbrace$.
Let $c_{policy}$ be a constant heuristic that estimates how likely a tactic
$t_i$ is to prove the goal.
We can now calculate of a prior policy of a child $a_i$ produced by the tactic
$t_i$ by:
\[PriorPolicy(a_i) = (1 - c_{policy})^{i} * c_{policy}\]

@@

We now concentrate on the definition a reasonable evaluation function for the
output of tactics. Theoretically, an interesting list of goals is one
that has short proofs for each of its goals, and we could estimate their 
lengths from previous proof searches. However, this
approach generates too much data and slows down the proof search. So, we only 
collect output of tactics tested during orthogonalization. We
declare a list of goals $l$ to be positive if it is has been produced by the
winner of an orthogonalization competition. 
The set of positive examples in 
$\mathit{Predictions}^{\mathit{goal\_list}}_k(l)$ is noted 
$\mathit{Positive}^{\mathit{goal\_list}}_k(l)$.
We chose $k=10$ as a default evaluation radius in our experiments.
And we evaluate a node $a$ through the list of goals $G(a)$ it
contains using the prior evaluation function:
\begin{align*}
\mathit{PriorEvaluation}_k (a) &= 
  \frac{card\ \mathit{Positive}^{\mathit{goal\_list}}_k(G(a)}{k}\\
\end{align*}

\subsection{ATP Integration}~\label{sec:atp}
General-purpose proof automation mechanisms which combine proof translation to
ATPs with machine learning (``hammers'') have become quite successful in
enhancing the automation level in proof assistants~\cite{hammers4qed}.
As external automated reasoning techniques sometimes outperform the combined
power of tactics, we would like to combine the \tactictoe search with
general purpose provers such as the ones found in \holyhammer for
\holfour~\cite{tgck-cpp15}.

Such a prover already exists in \holfour and is called \metis. It is already
recognized by the tactic selection mechanism in \tactictoe and can have its
premises predicted dynamically when in an abstracted form. Nevertheless,
we think that the performance of \tactictoe can be boosted by giving the ATP
\metis a special status. This arrangement consists of always predicting
premises for \metis, giving it a slightly higher timeout and
trying it first on each open goal. These modifications only induce an
linear overhead on the number nodes as a general purpose prover does not create
goal obligations.

The integration of an external prover is also possible. And
\eprover~\cite{eprover} is the one we experimented with.
To this end, we created a tactic which expects a number of premises as an 
argument
and calls \eprover on the translated first order problem. We then extract the
premises that were required in the \eprover
proof and reconstruct it inside \holfour with \metis. Giving a special status
to external
provers is essential as external ATP calls do not appear in human
proof scripts. For external ATPs, we use an even higher timeout (5 seconds) and 
a larger number of predicted premises (128 for \eprover) and also try them 
first on open 
goals. Since calls to ATPs are
computationally expensive, they are run in parallel and asynchronously. This
avoids slowing down \tactictoe's search loop. The number of asynchronous calls
to external provers that can be executed at the same time is limited by the
number of cores available to the user.

@@

Here are the different kind of feature we extract from terms:
\begin{itemize}
\item names of constants, including the logical operators,
\item type constructors present in the types of constants and variables,
\item subterms with all variables replaced by a single place holder $V$,
\item names of the variables.
\end{itemize} 
Goals (respectively theorems) are represented by a couple of a list of 
terms (assumptions) and a term (conclusion). So, we distinguish between
features of the assumptions and features of the conclusion by adding a
different tag to elements of each set. The union of all those 
features are considered the feature of a goal (respectively theorem). From 
that, we can construct features for a list of goals by
from the union of the features of each goal in the list.

@@

The first idea is to select tactic-goal pairs and theorems by their similarity 
with $c$ (i.e $\mathit{Predictions}^{\mathit{tactic}}_{500} (c)$ and
$\mathit{Predictions}^{\mathit{theorem}}_{500} (c)$).
However, this selection may not be adapted to later stages of
the proof where goals may have drifted significantly from $c$. In order to
anticipate the production of diverse goals, our prediction during preselection
takes dependencies between objects of the same dataset into account.
This dependency relation is assymetric. Through the relation, each object has
multiple children and at most one parent.
Once a dependency relation is established we can calculate a dependency score
for each object, which is the maximum of its similarity score 
and the similarity score of its parent. In the end, the 500 entries with 
highest dependency score are preselected in each dataset.

@@

\section{Abstraction}\label{sec:synthesis}
One of the great weakness of the previous version of \tactictoe is that
it could not create its own tactics. Indeed, sometimes no previous tactic is
adapted for a certain goal and creating a new tactic is necessary.
%Experiments in section, show how different degrees of tactic synthesis could
%improve the success rate of \tactictoe.
We present a way to create tactics with different arguments
by abstracting them and re-instantiating them using a predictor for that kind
of argument. In the spirit of the orthogonalization method, we will try to
create tactics that are more general but have the same effect as the
ungeneralized one. The generalized tactics are typically slower that their 
counterpart, but the added flexibility is worthwhile in practice. Moreover, 
since we impose a timeout on each tactic (0.05 seconds by default), very slow 
tactics fails and thus are not selected.

\paragraph{Abstraction of Tactic Arguments}
The first step is to abstract arguments of tactics by replacing them by a
place holder, creating a tactic with an unspecified argument.
Suppose we have a recorded tactic $t$. Since $t$ is a \sml code tree, we can
try to abstract any of the
\sml subterms. Let $u$ be a \sml subterm of $t$, and $h$ a placeholder $u$, we
can substitute $u$ by $h$ to create an abstracted tactic.
We note this intermediate tactic $t[h/u]$. By repeating the
process, multiple
arguments may be abstracted in the same tactic. Ultimately, the more
abstractions are performed the more general a tactic is, as many tactics
become instances of the same abstracted tactic. As a consequence, it becomes 
harder and harder to predict suitable arguments. In our experiments, we create 
one abstraction $\hat{t}$ for each tactic $t$ by abstracting all subterms of 
type theorem list in $t$.

\paragraph{Instantiation of an Abstracted Tactic}
An abstracted tactic is not immediately applicable to a goal, since it contains
unspecified arguments. To apply an abstracted tactic $\hat{t}$, we first need
to
instantiate the placeholders inside $\hat{t}$. Because it is difficult to
build a general predictor that effective on each type of argument, we manually
design different argument predictors for each type. Those predictors are
given a goal $g$ as input and returns the best chosen argument for this
goal.

Our default algorithm relies on argument predictions for theorems with a 
radius of 16. It uses the produced list 
$\mathit{Predictions}^{\mathit{tactic}}_{16}(g)$ to replace all the 
placeholders 
in $\hat{t}$. 
List of theorems is a very common type in \holfour proofs and theorems contain 
enough information to be predicted accurately.
We also have positive results with term abstraction and plan to include it soon 
in our default algorithm.

\paragraph{Selection of Abstracted Tactics}
As abstracted tactics do not appear in human proofs, we need to find a way to
predict them so that they can contribute to the \tactictoe proof search.
A straightforward idea is to try $\hat{t}$ before $t$ during the proof search.
However, this risks doing unnecessary work as they may have a similar effect.
Therefore, we would like to decide
beforehand if one is better than the other.
In fact, we can re-use the orthogonalization module to do this task for us.
We add $\hat{t}$ to the competition, initially giving it the coverage of $t$.
If $\hat{t}$
wins, it is associated with $g$ and is included in the database of tactic
features and thus can be predicted during proof search.
After many orthogonalization competitions, the coverage of $\hat{t}$ may exceed 
the coverage of
$t$. At
this point, the coverage of $\hat{t}$ is estimated on its own and not inherited
from
$t$ anymore.

@@

\subsection{Tuning \tactictoe}~\label{sec:tuning}

@@

%(more than 10 steps long).In all those experiments orthogonalization is

@@

\subsection{Full-scale experiment}~\label{sec:full_exp}

@@

The Table~\ref{theories} compares the success rates of re-proving for different
\holfour theories. \tactictoe outperforms \eprover on every
considered theory.
\eprover is more suited to deal with dense theories such as
\texttt{real} or \texttt{complex} where a lot of related theorems are available
and most proofs are usually completed by rewriting tactics. Thanks to its 
ability to re-use custom-built tactics, \tactictoe 
largely surpasses \eprover on theories relying on inductive terms and 
simplifications sets such as \texttt{arithmetic}, \texttt{list} 
and \texttt{f\_map}. Indeed, \tactictoe is able to recognize where and when to 
apply induction, a task at which ATPs are known to struggle with.
